Software architectures for workflow integration in science gateways

### 2016-04-21 paper.tex
Updates in Section 2. Modified command to add notes.

\definecolor{ao(english)}{rgb}{0.0, 0.5, 0.0}
\newcommand{\todo}[1]{\color{blue}\xspace[\emph{#1}]\xspace\color{black}}
\newcommand{\note}[1]{\color{ao(english)}\textbf{Note: }\textit{#1} \color{black}}

@@

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

@@

\item Workflow integration: consists in adding a new workflow to the
  system so that users can execute it. It is triggered by an
  administrator of the science gateway and it results in an interface,
  for instance a web form, where users can enter the parameters of the
  workflow to be executed. The interaction has two aspects:
  (\texttt{a$_1$}) the programs used in the workflow are installed on
  the infrastructure, which may or may not require administrative
  privileges on the infrastructure; (\texttt{a$_2$}) the workflow is
  configured in the science gateway so that it becomes available to
  users. Note that integrating a workflow is not the same process as
  integrating a workflow \emph{engine}.
\item Task control: operations to manage tasks on the infrastructure,
  including: authentication, submission, monitoring, termination,
  deletion, etc. Controlling tasks requires to deal with the
  heterogeneous batch managers and meta-schedulers that might be
  available on the infrastructure. When the infrastructure is a grid
  or a cloud, it is for instance implemented using libraries such as
  SAGA, DRMAA, OCCI and similar initiatives \todo{refs needed}.
\item Data control: operations to manage data on the infrastructure,
  such as: upload, download, deletion, browsing, replication, caching,
  etc. Data movements can be triggered by the user in the science
  gateway (\texttt{$c_1$}), to upload input data or download processed
  data. They can also be performed by the workflow engine
  (\texttt{$c_2$}), to transfer data across the infrastructure. The
  infrastructure might offer various data storage backends with
  heterogeneous interfaces. \note{this -c2- can be also trigerred by the 
  user} Tools and services such as JSAGA~\cite{reynaud2010uniform} or
  Data Avenue~\cite{hajnal2014data} can be used to homogeneize these 
  interfaces.
\item Workflow control: operations to execute a workflow with an
  engine, including: workflow submission, monitoring, termination,
  etc. Workflow control can be coarse-grained (a.k.a. black box) or
  fine-grained (white box). In a coarse-grained model, the various
  tasks created by a workflow execution are masked and the user only
  has a global view of the workflow execution. In a fine-grained
  model, user is exposed to the workflow topology, i.e. to the outputs
  of the individual tasks, their statuses and so on.
\item Sub-task control: operations used by tasks to submit sub-tasks
  on the infrastructure, including: submission, monitoring,
  termination, deletion, etc. Sub-task control is similar to
  interaction \texttt{b}, except that information about the parent
  task which submitted a sub-task is usually available and used for
  additional control. For instance, the parent task may wait for all
  its sub-tasks to finish before finishing, and conversely all the
  sub-tasks may be killed if the parent task is killed.
\item Pool-agent: specific to the architecture described
  in Section~\ref{sec:pool}. This is an interaction used when agents retrieve
  work from a central pool. It covers agent registration and
  de-registration to the pool, protocols to send work from the pool to
  the agent, mechanisms to update work status, and so on. 
\item Workflow conversion: translation from one workflow language to
  another one. This interaction may not be available or implementable
  for every workflow language. It has been developed mostly for
  well-structured and relatively simple workflow language such as between GWorkflowDL
  and Scufl~\cite{OLAB-09} and between the 4 languages that were involved in the
  SHIWA initiative. In SHIWA, workflow conversion is done through an
  intermediate representation, the Interoperable Workflow Intermediate
  Representation~\cite{plankensteiner-montagnat-etal:2011}, which
  allows to convert among $n$ workflow languages using $2n$
  interactions instead of $n^2$.
\end{enumerate}

@@

\begin{figure}
\def\svgwidth{\columnwidth}
\input{figures/CBRAIN-PSOM.pdf_tex}
\caption{Architecture of the CBRAIN-PSOM integration that illustrates
  the sub-tasking model (Figure extracted from~\cite{GLAT-16}). 1:
  User sends data and workflow execution request to storage server(s)
  and CBRAIN portal. 2: CBRAIN portal sends execution request to
  execution server on cluster. 3: Execution server transfers data from
  storage server(s). 4-5: Execution server starts PSOM master via task
  scheduler. 6: PSOM master submits PSOM workers to execution
  server. 7-8: Execution server starts PSOM workers through task
  scheduler. 9: PSOM master and workers execute workflow. 10.
  Execution server transfers results to storage server(s). \todo{Improve graphics}}
\label{fig:cbrain-psom-architecture}
\end{figure}

>>>>

\definecolor{ao(english)}{rgb}{0.0, 0.5, 0.0}
\newcommand{\todo}[1]{\color{blue}\xspace[\emph{#1}]\xspace\color{black}}
\newcommand{\note}[2]{\textbf{\Large{\color{blue}\footnote{{\color{blue}\textbf{#1:} \textit{#2}\color{black}}}}}}

@@

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnotes;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

@@

\item Workflow integration: consists in adding a new workflow to the
  system so that users can execute it. It is triggered by an
  administrator of the science gateway and it results in an interface,
  for instance a web form, where users can enter the parameters of the
  workflow to be executed. The interaction has two aspects:
  (\texttt{a$_1$}) the programs used in the workflow are installed on
  the infrastructure, which may or may not require administrative
  privileges on the infrastructure; (\texttt{a$_2$}) the workflow is
  configured in the science gateway so that it becomes available to
  users. Note that integrating a workflow is not the same process as
  integrating a workflow \emph{engine}.
\item Task control: operations to manage tasks on the infrastructure,
  including: authentication, submission, monitoring, termination,
  deletion, etc. Controlling tasks requires to deal with the
  heterogeneous batch managers and meta-schedulers that might be
  available on the infrastructure. When the infrastructure is a grid
  or a cloud, it is for instance implemented using libraries
  implementing standards such as SAGA~\cite{goodale2006saga},
  DRMAA~\cite{troger2012distributed} or OCCI~\cite{edmonds2012toward}.
\item Data control: operations to manage data on the infrastructure,
  such as: upload, download, deletion, browsing, replication, caching,
  etc. Data movements can be triggered by the user in the science
  gateway (\texttt{$c_1$}), to upload input data or download processed
  data. They can also be performed by the workflow engine
  (\texttt{$c_2$}), to transfer workflow data (inputs, outputs, or
  intermediate results) across the infrastructure so that tasks can
  use it. The infrastructure might offer various data storage backends
  with heterogeneous interfaces. \note{Marc-Etienne}{this -c2- can be also trigerred
    by the user}\note{Tristan}{No, it can't: this is by definition an interaction between the workflow engine and the infrastructure. It is used to transfer data to/from tasks. I reworded a bit, is it clearer now?} Tools and services such as
  JSAGA~\cite{reynaud2010uniform} or Data Avenue~\cite{hajnal2014data}
  can be used to homogeneize these interfaces.
\item Workflow control: operations to execute a workflow with an
  engine, including: workflow submission, monitoring, termination,
  etc. Workflow control can be coarse-grained (a.k.a. black box) or
  fine-grained (white box). In a coarse-grained model, the various
  tasks created by a workflow execution are masked and the user only
  has a global view of the workflow execution. In a fine-grained
  model, user is exposed to the workflow topology, i.e. to the outputs
  of the individual tasks, their statuses and so on.
\item Sub-task control: operations used by tasks to submit sub-tasks
  on the infrastructure, including: submission, monitoring,
  termination, deletion, etc. Sub-task control is similar to
  interaction \texttt{b}, except that information about the parent
  task which submitted a sub-task is usually available and used for
  additional control. For instance, the parent task may wait for all
  its sub-tasks to finish before finishing, and conversely all the
  sub-tasks may be killed if the parent task is killed.
\item Pool-agent: specific to the architecture described
  in Section~\ref{sec:pool}. This is an interaction used when agents retrieve
  work from a central pool. It covers agent registration and
  de-registration to the pool, protocols to send work from the pool to
  the agent, mechanisms to update work status, and so on. 
\item Workflow conversion: translation from one workflow language to
  another one. This interaction may not be available or implementable
  for every workflow language. It has been developed mostly for
  well-structured and relatively simple workflow language such as between GWorkflowDL
  and Scufl~\cite{OLAB-09} and between the 4 languages that were involved in the
  SHIWA initiative. In SHIWA, workflow conversion is done through an
  intermediate representation, the Interoperable Workflow Intermediate
  Representation~\cite{plankensteiner-montagnat-etal:2011}, which
  allows to convert among $n$ workflow languages using $2n$
  interactions instead of $n^2$.
\end{enumerate}

@@

\begin{figure}
\def\svgwidth{\columnwidth}
\input{figures/CBRAIN-PSOM.pdf_tex}
\caption{Architecture of the CBRAIN-PSOM integration that illustrates
  the sub-tasking model (Figure extracted from~\cite{GLAT-16}). 1:
  User sends data and workflow execution request to storage server(s)
  and CBRAIN portal. 2: CBRAIN portal sends execution request to
  execution server on cluster. 3: Execution server transfers data from
  storage server(s). 4-5: Execution server starts PSOM master via task
  scheduler. 6: PSOM master submits PSOM workers to execution
  server. 7-8: Execution server starts PSOM workers through task
  scheduler. 9: PSOM master and workers execute workflow. 10.
  Execution server transfers results to storage server(s).}
\label{fig:cbrain-psom-architecture}
\end{figure}
